\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-05-19 forwards-and-backwards.tex
.
figures
Update on Overleaf.

\comment{TODO: talk about $N \to \infty$ rather than $N \to \infty$.}

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    Recall that we have defined the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Normal
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although sometimes we omit the dependence of $\meanq$ and $\covq$ on $x$).
    This implies that if we define, for $f : \IR^d \to \IR$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } N \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    $$
    \DG^* f(x)
        =
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq_i) .
    $$
\end{definition}

@@

\begin{corollary}[Deterministic limits]
    \label{cor:nonlocal_pde_limits}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the limit is deterministic (i.e., $\theta / N \to \alpha = 0$),
    then any limit $\eta_t$ is a weak solution to the nonlocal partial differential equation
    \begin{equation} \label{eqn:nonlocal_pde}
        \partial_t \varphi_t(x)
        =
        r\left(x, \smooth{r} \varphi_t(x) \right)
        \DG^* \left[
            \varphi_t(\cdot)
            \gamma\big( x, \smooth{\gamma} \varphi_t(\cdot) \big)
        \right](x)
        +
        \varphi_t(x)
        F\left(x, \smooth{F} \varphi_t() \right)
        ,
    \end{equation}
    in the sense of Definition~\ref{defn:weak_solutions}.
    If equation \eqref{eqn:nonlocal_pde} has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges to that solution.
\end{corollary}

@@

Now that we have established what we can say about how population density changes with time,
we turn to results on ancestral lineages,
i.e., how genealogical ancestry can be traced back across the landscape.
Informally,
a \emph{lineage} $(L_t)_{t \ge 0}$
begun at spatial location $L_0 = x$
can be obtained by picking a focal individual uniformly at $x$,
and then for each amount of time $t$ in the past,
setting $L_t$ to the spatial location of the individual from whom
the focal individual inherits at that time in the past.
Since in our model individuals have only one parent, this is unambiguous.
Although we did not explicitly retain such information,
it is clear that for finite $N$
one could construct the distribution of a lineage $(L_t)_{t=0}^T$
given the history of the population $(\eta^N_t)_{t = 0}^T$,
for each starting location to which $\eta^N_T$ assigns positive mass.
It is less clear, however, how to formally retain such information in the limit.
The \emph{lookdown construction} in Section~\ref{sec:lookdown}
provides just such a construction,
where lineages are retained in the $N \to \infty$ limit.
Roughly speaking,
this works by assigning to each particle a unique ``level''
that functions as a label and thus allows reconstruction of lineages,
but also (in some sense) orders individuals by eventual reproductive output,
allowing a formal limit.
See~\ref{etheridge/kurtz:2018} for an introduction to these ideas.

@@

\begin{figure}
    \begin{center}
        FIGURE
    \end{center}
    \caption{
        Two panels, showing simulated 1D populations under the PME, comparing the simulated
        profile to the analytical solution;
        $\theta/N$ not small on left; small on right.
        (Note: noisier wave should move slower, we'll see this.)
        \label{fig:pme_waves}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
        FIGURE
    \end{center}
    \caption{
        Six panels: one pair for FKPP, one for PME, and one for Allen-Cahn.
        Each pair shows (top) lineage traced back in wavefront,
        and (bottom) stationary distribution of lineage location,
        compared to analytical solution for PME.
        \label{fig:pme_vs_fkpp}
    }
\end{figure}

@@

%% %% %% %% %% %% %% %% %% %% %% %%
\subsection{Construction tracking individual lines of descent}

@@

@@

\begin{proposition}[Line of descent construction]
Let $\beta(x, \eta) = \meanq + 2 \covq \grad \log r(x, \eta)$
and $K$ such that $\covq = K K^T$.
Associate with each label $a \in \labelspace$
an independent $d$-dimensional Brownian motion $W_a$
and an independent Poisson process $R_a$ on $[0, \infty)^2$
with Lebesgue mean measure,
as before with points ordered in some way.
Given $\eta_0 \in \measures$,
let $(x_i, u_i)$ be the points of a Poisson process
with mean measure on $\IR^d \times [0, \infty)$ a product of $\eta_0$ and Lebesgue measure.
For each point begin a line of descent
with label $i$, location $X_i(0) = x_i$, level $U_a(0) = u_i$, and birth time $\tau_i = 0$.
Suppose that the spatial locations and level of each line of descent $a$
behave, for $\tau_a \le t < \sigma_a$ as, as
\begin{align*}
X_a(t)
    &=
    X_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \beta(x, \eta(s)) ds
        +
        K dW_a(s)
    \bigg)
    ds  \\
U_a(t)
    &=
    U_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \alpha \gamma(X_a(s),\eta(s))
        r(x,\eta(s)) U_a(s)^2
\\ &\qquad \qquad \qquad {}   
        -
        \left\{
            \gamma(X_a(s),\eta(s)) \DG r(X_a(s),\eta(s))
            + F(X_a(s), \eta(s))
        \right\}
        U_a(s)
    \bigg)
    ds .
\end{align*}
Each point in the $R_a$ denotes a potential birth time for $a$:
if the $j^\text{th}$ point in $R_a$ is $(t, v)$, with $t < \sigma_a - \tau_a$,
then a new line of descent with label $a \concat j$ is produced,
with birth time $\tau_{a \concat j} = \tau_a + t$,
location $X_{a \concat j}(\tau_a + t) = X_a(tau_a + t)$, and level
$$
    U_{a \concat j}(tau_a + t) = U_a(tau_a + t)
    + \frac{v}{ 2 \alpha \gamma(X_a(\tau_a + t), \eta_{tau_a + t}) r(X_a(\tau_a + t), \eta_{\tau_a + t}) } .
$$
For each $a$, let $\sigma_a = \inf\{t \ge 0: U_a(t) < \infty\}$.
Then, the processes defined by
\begin{align*}
    \eta_t = \lim_{N \to \infty} \frac{1}{N} \sum_{a : \tau_a \le t < \sigma_a} \delta_{X_a(t)}
    \qquad \text{and} \qquad
    \lp_t = \sum_{a : \tau_a \le t < \sigma_a} \delta_{(X_a(t), U_a(t))} 
\end{align*}
agree in distribution with those of Definitions~\ref{defn:mgale_construction} and~\ref{defn:lookdown_mgale},
respectively.
\end{proposition}

@@

>>>>

\comment{TODO: talk about $N \to \infty$ rather than $\theta \to \infty$.}

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    Recall that we have defined the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Normal
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although sometimes we omit the dependence of $\meanq$ and $\covq$ on $x$).
    This implies that if we define, for $f : \IR^d \to \IR$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq(x)_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq(x)_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } N \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    $$
    \DG^* f(x)
        =
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq(x)_i) .
    $$
\end{definition}

@@

\begin{corollary}[Deterministic limits]
    \label{cor:nonlocal_pde_limits}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the limit is deterministic (i.e., $\theta / N \to \alpha = 0$),
    then any limit $\eta_t$ is a weak solution to the nonlocal partial differential equation
    \begin{equation} \label{eqn:nonlocal_pde}
        \partial_t \varphi_t(x)
        =
        r\left(x, \smooth{r} \varphi_t(x) \right)
        \DG^* \left[
            \varphi_t(\cdot)
            \gamma\big( x, \smooth{\gamma} \varphi_t(\cdot) \big)
        \right](x)
        +
        \varphi_t(x)
        F\left(x, \smooth{F} \varphi_t(x) \right)
        ,
    \end{equation}
    in the sense of Definition~\ref{defn:weak_solutions}.
    If equation \eqref{eqn:nonlocal_pde} has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges to that solution.
\end{corollary}

@@

Now that we have established what we can say about how population density changes with time,
we turn to results on ancestral lineages,
i.e., how genealogical ancestry can be traced back across the landscape.
Informally,
a \emph{lineage} $(L_t)_{t \ge 0}$
begun at spatial location $L_0 = x$
can be obtained by picking a focal individual uniformly at $x$,
and then for each amount of time $t$ in the past,
setting $L_t$ to the spatial location of the individual from whom
the focal individual inherits at that time in the past.
Since in our model individuals have only one parent, this is unambiguous.
Although we did not explicitly retain such information,
it is clear that for finite $N$
one could construct the distribution of a lineage $(L_t)_{t=0}^T$
given the history of the population $(\eta^N_t)_{t = 0}^T$,
for each starting location to which $\eta^N_T$ assigns positive mass.
It is less clear, however, how to formally retain such information in the limit.
The \emph{lookdown construction} in Section~\ref{sec:lookdown}
provides just such a construction,
where lineages are retained in the $N \to \infty$ limit.
Roughly speaking,
this works by assigning to each particle a unique ``level''
that functions as a label and thus allows reconstruction of lineages,
but also (in some sense) orders individuals by eventual reproductive output,
allowing a formal limit.
See~\citet{etheridge/kurtz:2018} for an introduction to these ideas.

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figures/ex2a/pme_123.steps}
        \includegraphics[width=0.45\textwidth]{figures/ex2b/pme_123.steps}
    \end{center}
    \caption{
        Two panels, showing simulated 1D populations under the PME, comparing the simulated
        profile to the analytical solution;
        $\theta/N$ not small on left; small on right.
        (Note: noisier wave should move slower, we'll see this.)
        \label{fig:pme_waves}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
        \includegraphics{figures/ex3_fkpp/fkpp_123.lineages}
        \includegraphics{figures/ex3_fkpp/fkpp_123.lineagehist}
        \includegraphics{figures/ex3_pme/pme_123.lineages}
        \includegraphics{figures/ex3_pme/pme_123.lineagehist}
        \includegraphics{figures/ex3_allen-cahn/allen-cahn_123.lineages}
        \includegraphics{figures/ex3_allen-cahn/allen-cahn_123.lineagehist}
    \end{center}
    \caption{
        Six panels: one pair for FKPP, one for PME, and one for Allen-Cahn.
        Each pair shows (top) lineage traced back in wavefront,
        and (bottom) stationary distribution of lineage location,
        compared to analytical solution for PME.
        \label{fig:pme_vs_fkpp}
    }
\end{figure}

@@

%% %% %% %% %% %% %% %% %% %% %% %%
\subsection{Construction tracking individual lines of descent} \label{sec: individual lines of descent}

@@

\subsubsection{Evolution of levels along a whole line of descent}
We now write down expressions
that give the spatial and level dynamic
of a whole line of descent.
In particular, our description is consistent
for any $N \in \mathbb{N}$.

Recall that
the particle with label $(a_0,a_1,..,a_k)$
is the offspring
with a new level
from the $a_k$-th potential birth
of $(a_0,...,a_{k-1})$,
who is again the offspring
with a new level from
the $a_{k-1}$-th potential birth
of $(a_0,...,a_{k-2})$, 
etc..
In particular, 
the particle with label $(a_0)$
is the direct descendant of
the $a_0$-th original ancestor
of the whole population, i.e.
it has never experienced
a jump in level throughout
its ancestry.

For simplicity, for $a=(a_0,..,a_k)$,
we denote $[a]_{i}:=(a_0,..,a_i)$
to be the $i$-th generation 
ancestor of the lineage $a$.
In particular, if at time $t$,
if the ancestral lineage $[a]_t$
is at its $i(t)$-th generation,
then $[a]_t = [a]_{i(t)}$.

We also write 
$(s_{[a]_{i}}, v_{[a]_{i}}, y_{[a]_{i}}, \kappa_{[a]_{i}})
\in \Pi_{[a]_{i-1}}$
to be the reproduction event
that creates $[a]_{i}$ from $[a]_{i-1}$,
and similarly,
$(s_{[a]_{i} \concat j}, v_{[a]_{i} \concat j}, y_{[a]_{i} \concat j}, \kappa_{[a]_{i} \concat j})
\in \Pi_{[a]_{i-1}}$
to be the reproduction event 
that creates $[a]_{i} \concat j$
from $[a]_{i-1}$.

We first write down an expression
for the level of particle $a=(a_0,a_1,..,a_k)$ at time $t$.
Since we ensure that
the level of a particle
only jumps when a reproduction event
creates a new label,
the level of particle $a=(a_0,a_1,..,a_k)$ is determined
by $k$ many jumps,
and subsequent continuous evolutions
given by Equation \eqref{eqn:U_line_of_descent}.

Under this notation, for $t> 0$,
\begin{equation}
\label{eq: level_across_lineage}
\begin{aligned}
&U^{N}_{[a]_{k}}(t)- U^{N}_{[a]_{0}}(0)\\
=& \sum_{i=1}^{k} \int_{\tau_{[a]_{i-1}}}^{\tau_{[a]_{i}}}
\left(
        \frac{\theta}{N} \gamma(X_{[a]_{i-1}}(s),\eta_s)
        \int_{\IR^d} r(z,\eta_s) q(X_{[a]_{i-1}}(s),dz) U_{[a]_{i-1}}(s)^2
        -
        b(X_{[a]_{i-1}}(s),\eta_s) U_{[a]_{i-1}}(s)
    \right)
    ds\\
&+\int_{\tau_{[a]_{i}}}^{t}
\left(
        \frac{\theta}{N} \gamma(X_{[a]_{i}}(s),\eta_s)
        \int_{\IR^d} r(z,\eta_s) q(X_{[a]_{i}}(s),dz) U_{[a]_{i}}(s)^2
        -
        b(X_{[a]_{i}}(s),\eta_s) U_{[a]_{i}}(s)
    \right)
    ds\\
&+\sum_{i=1}^{k}
    \left( \ell(X_{[a]_{i-1}}(\tau_{[a]_{i}}-), y_{[a]_{i}}, \eta(\tau_{[a]_{i}}-), v_{[a]_{i}})
    \right)
\end{aligned}    
\end{equation}
where the first two terms correspond to
sum of the evolution along the same generation,
and the second term corresponds
to the sum of jumps in levels
when a new label is introduced.

To preserve consistency 
as we send $N \to \infty$
in our scaling limit,
we allow $U_{[a]_t}(t)$ to hit $\infty$.
To identify dead lines of descent,
for a label $a$,
we define
$\sigma_{a}^{N}:= \inf\{t \geq 0 : U_{[a]_t}(t) > N\}$.
In the $N$-th scaling regime,
we proclaim the individual
$a$ to be dead
at time $t$
when $t > \sigma_{a}^{N}$

Note that in particular,
if $U_{[a]_i}(t)>N$
at some time $t$
in the $i$-th generation,
then $U_{[a]_j}(t)>N$
for all $j \geq i$
and
$\sigma_{[a]_j}^{N}=\sigma_{[a]_i}^{N}$.
In words, all descendants of dead
particles are dead by default.

Similarly, we generalise the notion of death above
to the infinity case by considering 
$\sigma_{[a]_t}^{\infty}:=\inf\{t \geq 0 : U_{[a]_t}(t) =\infty\}$.

\subsubsection{Spatial jumps along a whole line of descent}

We now write down an expression for the
spatial location of $[a]_t$ at time $t$.
Note that in the spatial case,
jumps occur along the same generation
whenever $\kappa = 0$ in successful birth events,
and occur across two generations
whenever $\kappa = 1$ in successful birth
events.

To illustrate this, consider the spatial location of $(a_0,a_1)=(2,3)$ with birth time $\tau_{(2,3)}$ at some time $t>\tau_{(2,3)}$.
If $U_{(2,1)}(\tau_{(2,1)}) < N$,
the first birth event of the ancestor
with label $(2)$
is successful.
If $\kappa = 0$ in this birth event,
the parent and offspring swap labels,
resulting in
a spatial jump for the label $(2)$.
On the contrary, 
if $U_{(2,2)}(\tau_{(2,2)}) \geq N$
or $\kappa = 1$
in the second birth event of $(2)$,
the parent does not give birth
or retains its label,
so there is no spatial jump.

On the other hand, 
if $\kappa=1$ when $(2,3)$ is born,
the new label is given to
the offspring
with new spatial location,
resulting in a jump,
and vice versa.

After $(2,3)$ is born,
whenever the particle with label
$(2,3)$ gives birth, 
if $\kappa = 0$, the label $(2,3)$
is adopted by the offspring
which results in a spatial jump
in the position of $(2,3)$
again.

Under this notation, for $t>\tau_{[a]_k}$,

\begin{equation}
\label{eq: space_across_lineage}
\begin{aligned}
&X^{N}_{[a]_{k}}(t)- X^{N}_{[a]_{0}}(0)\\
=& \sum_{i=1}^{k}
\sum_{j=1}^{a_i-1}
(1-\kappa_{[a]_{i-1}\concat j}) y_{[a]_{i-1}\concat j} \ind_{U_{[a]_{i-1}\concat j}(\tau_{[a]_{i-1}\concat j}) < N}\\
&+ \sum_{i=1}^{k}
\kappa_{[a]_i} y_{[a]_i} \ind_{U_{[a]_i}(\tau_{[a]_i}) < N}\\
&+\int_{\tau_{[a]_k}}^{t}
(1-\kappa)y
\ind_{ \{
        U_{[a]_k}(\tau_{[a]_k}+s)
        +\ell(X_{[a]_k}(\tau_{[a]_k}+s),y,\eta_{[a]_k}(\tau_{[a]_k}+s),v) < N
        \}
      } 
\Pi_{[a]_k}(dsdvdyd\kappa),
\end{aligned}    
\end{equation}
where the first term
corresponds to 
the sum of $a_{i}-1$ many potential jumps
in the $(i-1)-th$ generation;
the second term corresponds to 
the sum of $k$ many potential jumps
when the $k$-th generation is introduced;
and the last term
corresponds to sum of 
potential jumps that occured in
the current generation from 
time $\tau_{[a]_k}$
to current time $t >\tau_{[a]_k}$.

@@

\begin{proposition}[Line of descent construction]
Let $\beta(x, \eta) = \meanq + 2 \covq \grad \log r(x, \eta)$
and $K$ such that $\covq = K K^T$.
Associate with each label $a \in \labelspace$
an independent $d$-dimensional Brownian motion $W_a$
and an independent Poisson process $R_a$ on $[0, \infty)^2$
with Lebesgue mean measure,
as before with points ordered in some way.
Given $\eta_0 \in \measures$,
let $(x_i, u_i)$ be the points of a Poisson process
with mean measure on $\IR^d \times [0, \infty)$ a product of $\eta_0$ and Lebesgue measure.
For each point begin a line of descent
with label $i$, location $X_i(0) = x_i$, level $U_a(0) = u_i$, and birth time $\tau_i = 0$.
Suppose that the spatial locations and level of each line of descent $a$
behave, for $\tau_a \le t < \sigma_a$ as, as
\begin{align*}
X_a(t)
    &=
    X_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \beta(x, \eta(s)) ds
        +
        K dW_a(s)
    \bigg)
    ds  \\
U_a(t)
    &=
    U_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \alpha \gamma(X_a(s),\eta(s))
        r(x,\eta(s)) U_a(s)^2
\\ &\qquad \qquad \qquad {}   
        -
        \left\{
            \gamma(X_a(s),\eta(s)) \DG r(X_a(s),\eta(s))
            + F(X_a(s), \eta(s))
        \right\}
        U_a(s)
    \bigg)
    ds .
\end{align*}
Each point in the $R_a$ denotes a potential birth time for $a$:
if the $j^\text{th}$ point in $R_a$ is $(t, v)$, with $t < \sigma_a - \tau_a$,
then a new line of descent with label $a \concat j$ is produced,
with birth time $\tau_{a \concat j} = \tau_a + t$,
location $X_{a \concat j}(\tau_a + t) = X_a(tau_a + t)$, and level
$$
    U_{a \concat j}(\tau_a + t) = U_a(\tau_a + t)
    + \frac{v}{ 2 \alpha \gamma(X_a(\tau_a + t), \eta_{\tau_a + t}) r(X_a(\tau_a + t), \eta_{\tau_a + t}) } .
$$
For each $a$, let $\sigma_a = \inf\{t \ge 0: U_a(t) < \infty\}$.
Then, the processes defined by
\begin{align*}
    \eta_t = \lim_{N \to \infty} \frac{1}{N} \sum_{a : \tau_a \le t < \sigma_a} \delta_{X_a(t)}
    \qquad \text{and} \qquad
    \lp_t = \sum_{a : \tau_a \le t < \sigma_a} \delta_{(X_a(t), U_a(t))} 
\end{align*}
agree in distribution with those of Definitions~\ref{defn:mgale_construction} and~\ref{defn:lookdown_mgale},
respectively.
\end{proposition}

@@

In this section,
we will prove tightness
of the spatial and level trajectory
of a line of descent
(constructed in \S \ref{sec: individual lines of descent})
as a forward-in-time process.

In Appendix \ref{sec: Topologies on Lookdown},
we will prove that
tightness of the whole lookdown process
holds given
tightness of 
all ancestral lineages
in a countable product topology.

To prove tightness of the 
spatial and level locations
of a typical line of descent,
we will show that the
jumps in
spatial and level locations
of the line of descent
in finite time
can be controlled.

Recall that the lines of descents
track chains of descendant
individuals forwards through time following the \textit{levels}.

\subsubsection{Tightness of Spatial Trajectories}
Before proving tightness of our spatial trajectories, 
note that for a fixed label
$a=(a_0,a_1,..,a_k)$,
the spatial evolution of 
$X^N_{a}(t)$ for $t \geq \tau^N_{a}$
is determined
purely by the last term 
in Equation \eqref{eq: space_across_lineage}.

As a result, the spatial jumps are 
driven by 
an underlying Poisson point process. 
It is therefore sufficient
to prove that square of 
these increments are bounded
in small time scales.

\begin{lemma} \label{lem: tightness of individual spatial trajectory}
For fixed $a=(a_0,a_1,..,a_k)$,
$\{(X^{N}_{a}(t))_{t \geq \tau^{N}_a}: N \in \mathbb{N},  \}$ is relatively compact in 
$D_{\mathbb{R}^d}[0,\infty )$ where $\mathbb{R}^d$ is equipped with the standard Euclidean metric. 
\end{lemma}
\begin{proof}
First,
note that the birth time
of $a$ at the $N$-th scaling
regime, denoted as
$\tau^{N}_a$,
is identical in distribution
for all $N > 0$.
This is because the birth time
of $a=(a_0,a_1,...,a_k)$
comes after $a_1$ births
in the 1st generation,
followed by $a_2$ births
in the 2nd generation, etc..
Therefore $\tau^N_{a} \sim Gamma(\sum_{i=1}^{k}a_i,1)$.

Now we prove
convergence in distribution
of the birth locations $X^N_{a}(\tau^N_{a})$.
From Equation \eqref{eq: space_across_lineage},
we have
\begin{equation}
\begin{aligned}
&X^{N}_{[a]_{k}}(\tau^N_{a})- X^{N}_{[a]_{0}}(0)\\
=& \sum_{i=1}^{k}
\sum_{j=1}^{a_i-1}
(1-\kappa_{[a]_{i-1}\concat j}) y_{[a]_{i-1}\concat j} \ind_{U^N_{[a]_{i-1}\concat j}(\tau_{[a]_{i-1}\concat j}) < N}\\
&+ \sum_{i=1}^{k}
\kappa_{[a]_i} y_{[a]_i} \ind_{U^N_{[a]_i}(\tau_{[a]_i}) < N}.
\end{aligned}    
\end{equation}

Since $(s_{[a]_{i} \concat j}, v_{[a]_{i} \concat j}, y_{[a]_{i} \concat j}, \kappa_{[a]_{i} \concat j})
\in \Pi_{[a]_{i-1}}$
are all iid random variables 
for each $i$ and $j$,
and that the sum above is finite,
it suffices to show that
the quantities 
\begin{equation}
\mathbb{P}(U^N_{[a]_{i-1} \concat j}(\tau_{[a]_{i-1} \concat j}) < N)    
\end{equation}
converges
for all $i=1,..,k$ and $j=1,..,a_i$.

Writing out
$U^N_{[a]_{i-1} \concat j}(\tau_{[a]_{i-1} \concat j})$,
we have
\begin{equation}
\begin{aligned}
&\mathbb{P}(U^N_{[a]_{i-1} \concat j}(\tau_{[a]_{i-1} \concat j}) < N) \\
=&\mathbb{P}(U^N_{[a]_{i-1} }(\tau_{[a]_{i-1} \concat j})
    +
    \ell(
    X^N_{[a]_{i-1} }(\tau_{[a]_{i-1} \concat j}-),
    y_{[a]_{i-1} \concat j},
    \eta^{N}(\tau_{[a]_{i-1} \concat j}-),
    v_{[a]_{i-1} \concat j})
    < N)\\
=&\mathbb{P}\left(
    \frac{
        v_{[a]_{i-1} \concat j}
    }{
        2 N^{-1} \theta \gamma(x, \eta) r(x + y, \eta)
    }
    < N-U^N_{[a]_{i-1} }(\tau_{[a]_{i-1} \concat j})
    \right)\\
=&\mathbb{P}\left(
            v^N_{[a]_{i-1} \concat j}
    < (2 \theta \gamma(x, \eta) r(x + y, \eta)-  2 N^{-1} \theta \gamma(x, \eta) r(x + y, \eta)U^N_{[a]_{i-1} }(\tau_{[a]_{i-1} \concat j}))
    \right)\\
=& \frac{(2 \theta \gamma(x, \eta) r(x + y, \eta)-  2 N^{-1} \theta \gamma(x, \eta) r(x + y, \eta)U^N_{[a]_{i-1} }(\tau_{[a]_{i-1} \concat j}))}{N}\\
\to & 2\alpha \gamma(x, \eta) r(x + y, \eta).
\end{aligned}    
\end{equation}

We write $S_{t,r}:=(t,t+r  ]\times [0,\infty )\times [0,1]\times 
\{0,1\}$ to be the region for the Poisson measure driving $(X^{\varepsilon,\theta, \lambda}_{\mathfrak{u}}(t))$ between time $t$ and $(t+r)$ and write $\tilde\Pi_{\tau_u}(ds,dv,dw,d\kappa):=\Pi_{\tau_u}(ds,dv,dw, d\kappa) - ds\times dv\times dw \times (\frac{1}{2}\delta_{0}+\frac{1}{2}\delta_{1})d\kappa$ be the compensated Poisson point process of $\Pi_{\tau_u}$. 
For $t\geq\tau_\mathfrak{u}$, we can compute
\small
\begin{eqnarray*}
&&\mathbb{E}[|X^{\varepsilon,\theta, \lambda}_{\tau_{\mathfrak{u}}}(t+r )-X^{\varepsilon,\theta, \lambda}_{\tau_{\mathfrak{u}}}(t)|^2|{\cal F}_t^{
\eta}]\\
&&= \mathbb{E}\left[\left|\int_{S_{t,r}}{\bf 1}_{\{U_{\tau_{\mathfrak{u}}}(s)+\frac v{2\lambda^{-1}\theta \gamma^{\mathfrak{m}}_\theta(X^{\varepsilon,\theta, \lambda}_{
\tau_{\mathfrak{u}}}(s-),\overline{\eta}^{\varepsilon,\theta, \lambda}(s-))}<\lambda \}}\kappa 
H(X^{\varepsilon,\theta, \lambda}_{\tau_{\mathfrak{u}}}(s-),\overline{\eta}^{\varepsilon,\theta, \lambda}(s-),w)\Pi_{\tau_{\mathfrak{u}}}(ds,dv,dw,d\kappa )\right|^2\Bigg|{\cal F}_t^{\eta}\right]\\
&&\leq 2\mathbb{E}\left[\left|\int_{S_{t,r}}{\bf 1}_{\{v < 2 \theta \gamma^{\mathfrak{m}}_\theta(X^{\varepsilon,\theta, \lambda}_{
\tau}(s-),\overline{\eta}^{\varepsilon,\theta, \lambda}(s-))\}}{\bf 1}_{\kappa =1}
H(X^{\varepsilon,\theta, \lambda}_{\tau}(s-),\overline{\eta}^{\varepsilon,\theta, \lambda}(s-),w)\tilde\Pi_{\tau_{\mathfrak{u}}}(ds,dv,dw,d\kappa )\right|^2\Bigg|{\cal F}_t^{\eta}\right]\\
&&\quad +2\mathbb{E}\left[\left|\int_{t}^{t+r}\int_{\{v < 2 \theta \gamma^{\mathfrak{m}}_\theta(X^{\varepsilon,\theta, \lambda}_{
\tau_{\mathfrak{u}}}(s-),\overline{\eta}^{\varepsilon,\theta, \lambda}(s-))\}}\frac{1}{2}\left| \int_{\mathbb{R}^d} (y-X^{\varepsilon,\theta, \lambda}_{\tau}(s-)) q^{\mathfrak{m}}_{\theta, \varepsilon}(X^{\varepsilon,\theta, \lambda}_{\tau}(s-),\overline{\eta}^{\varepsilon,\theta, \lambda}(s-),dy) \right|
dvds \right|^2\Bigg|{\cal F}_t^{\eta}\right]\\
&&\leq 2\mathbb{E}\left[\int_{t}^{(t+r)   }\theta (1+C_G)\int_{{\mathbb{R}}^d}|y-X^{\varepsilon,\theta, \lambda}_{
\tau_{\mathfrak{u}}}(s)|^2q^{\mathfrak{m}}_{\varepsilon,\theta}(X^{\varepsilon,\theta, \lambda}_{\tau_{\mathfrak{u}}}(s),\bar{\eta}^{\varepsilon,\theta,\lambda}(s-),dy)ds\Bigg|{\cal F}_t^{\eta}\right]\\
&&\quad +2\mathbb{E}\left[\left|\int_{t}^{(t+r)    } \theta (1+C_G)\left|\int_{{\mathbb{R}}^d}
(y-X^{\varepsilon,\theta, \lambda}_{\tau_{\mathfrak{u}}}(s))q^{\mathfrak{m}}_{\varepsilon,\theta}(X^{\varepsilon,\theta, \lambda}_{\tau_{\mathfrak{u}}}(s),
\overline{\eta}^{\varepsilon,\theta, \lambda}(s),dy)\right|ds\right|^2\Bigg|{\cal F}_t^{\eta}\right]\\
&&\leq 2r \overline{\beta}(1+C_G)+2(r \overline{\alpha}(1+C_G))^2.\end{eqnarray*}
\normalsize

The first inequality uses the fact that $(A+B)^2 \leq 2A^2+2B^2$ as well as the definition of the transition kernel $H$. The second inequality holds because birth events occur at rate $2 \theta \gamma^{\mathfrak{m}}_{\theta}$ bounded above by $2\theta (1+C_G)$. The third inequality holds given the assumptions on the dispersal kernel.

Finally, since $\overline{\eta}_0$ has compact support, we can take $t=\tau_u$ and apply Markov inequality to get that for every $\delta>0$, there exists $N_{\delta}>0$ such that
\begin{equation}\label{eq: Compact Containment Condition for individual spatial motion}
\sup_{\varepsilon, \theta, \lambda}\sup_{T>0}\mathbb{P}\left[|X^{\varepsilon,\theta, \lambda}_{\tau_{\mathfrak{u}}}(\tau_{\mathfrak{u}}+T)-X^{\varepsilon,\theta, \lambda}_{\tau_{\mathfrak{u}}}(\tau_{\mathfrak{u}})|>N_{\delta}\right]\leq   \frac{2r \overline{\beta}(1+C_G)+2(r \overline{\alpha}(1+C_G))^2}{N_{\delta}^2}  < \delta,
\end{equation}
giving us compact containment condition. Relative compactness thus follows by Theorem 3.8.6.
of \cite{ethier/kurtz:1986}.
\end{proof}
\begin{remark}
Note that under the assumptions laid out in Theorem \ref{teo: Non-local interaction Convergence} and Theorem \ref{teo: Local Interaction Convergence}, $r \equiv 1$, so 
$$\gamma^{\mathfrak{m}}_{\varepsilon,\theta}(x, \overline{\eta})=\tilde\gamma_{\theta}(x, \varrho_{\varepsilon}*\overline{\eta})= 1+\frac{G(\varrho_{\varepsilon}*\overline{\eta})}{\theta} \leq (1+C_G)$$
for all $\theta > 1$, and 
$$q^{\mathfrak{m}}_{\varepsilon,\theta,\lambda}(x,\overline{\eta},dy)=q_{\theta}(x,dy).$$
In particular, we have the inequalities
\begin{align}
   \sup_{\varepsilon, \theta}\limsup_{x, \overline{\eta}} \theta \left| \int_{\mathbb{R}^d} (y-x) q^{\mathfrak{m}}_{ \varepsilon, \theta}(x,\overline{\eta},dy) \right| =& \sup_{\theta}\limsup_{x} \theta \left| \int_{\mathbb{R}^d} (y-x) q_{\theta}(x,dy) \right| =  0,\\
   \sup_{\varepsilon, \theta}\limsup_{x, \overline{\eta}} \theta \int_{\mathbb{R}^d} |y-x|^2 q^{\mathfrak{m}}_{ \varepsilon, \theta}(x,\overline{\eta},dy)  =& \sup_{\theta}\limsup_{x} \theta \int_{\mathbb{R}^d} |y-x|^2 q_{\theta}(x,dy) =  1,
\end{align}
as the last term can be expressed as $\theta \mathbb{E}_{x}[(B_{1/\theta}-x)^2]=1$. As a result, Lemma \ref{lem: tightness of individual spatial trajectory} holds for the setting in Theorem \ref{teo: Non-local interaction Convergence} and Theorem \ref{teo: Local Interaction Convergence}.
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Tightness of Level Processes}
Proving tightness of the level process is slightly trickier as the levels of many particles will drift towards infinity in finite time. These particles however will not contribute to the metric associated with the vague topology and thus we will retain relative compactness. The main argument relies on the observation that particles with levels lower than some threshold $\overline{U}_{N, \delta}$  have regulated level increments of order $\delta$ over any small time interval of length $r$ given large enough $\lambda$. As a result, we can evaluate our level process against test functions $(g_n)_{n=1,..,N}$ with cut off smaller than $\overline{U}_{N, \delta}$.
